R and Python clients for GBIF species occurrence data

### 2016-03-11 manuscript.tex
converted ms to discuss python in addition to R client

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rgbif: R client for working with GBIF species occurrence data},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{rgbif: R client for working with GBIF species occurrence data}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
    \author[boettig]{Carl Boettiger}
   \ead{carl(at)ropensci.org}

@@

Perhaps the most fundamental element in many fields of ecology is the
individual. The number of individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while others
can easily take advantage of existing data. In fact, some ecology fields
are built largely on existing data, e.g., macro-ecology (Brown, 1995;
Beck et al., 2012).

@@

There are many online services that collect and maintain specimen
records. However, Global Biodiversity Information Facility (hereafter,
GBIF, \url{http://www.gbif.org}) is the largest collection of
biodiversity records globally, currently with 640 million records, 1.6
million taxa, 15,000 datasets from 780 publishers (current as of
2015-11-25). Many large biodiversity warehouses such as iNaturalist
(\url{http://www.inaturalist.org}), VertNet (\url{http://vertnet.org}),
and USGS's Biodiversity Information Serving Our Nation (BISON;
\url{http://bison.usgs.ornl.gov}) all feed into GBIF.

Herein, we describe the \texttt{rgbif} software library (Chamberlain et
al.) for working with GBIF data in the R programming environment (R Core
Team, 2014). R is a widely used language in academia, and in non-profit
and private sectors. Importantly, R makes it easy to do all of the steps
of the research process, including data management, data manipulation
and cleaning, statistics, and visualization. Thus, an R client for
getting GBIF data is a powerful tool to facilitate reproducible
research.

\section{The rgbif package}\label{the-rgbif-package}

@@

\subsection{Package interface}\label{package-interface}

\texttt{rgbif} is designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, taxonomic names, occurrences, and maps. We also include
functions to interface with the OAI-PMH GBIF service; only dataset
information is available vis this service, however. We ignore maps in
\texttt{rgbif} as it is concerned with generating maps primarily for web
applications. \texttt{rgbif} has a suite of functions dealing with each
of registry, taxonomic names, and occurrences - we'll go through each in
turn describing design and example usage.

\subsection{GBIF feedback loop}\label{gbif-feedback-loop}

With each request \texttt{rgbif} makes to GBIF's API, we send request
headers that tell GBIF that the request is coming from \texttt{rgbif},
including what version of the package. This helps GBIF know what
proportion of requests are coming from this package, and therefore R, as
most requests likely will come from \texttt{rgbif}; this information is
helpful for them in thinking about how people are using GBIF data.

@@

And dataset information in general is available via the OAI-PMH service,
functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
\NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
\CommentTok{#>  [1] "SDNHM Birds Collection"                                                }
\CommentTok{#>  [2] "CM Birds Collection"                                                   }
\CommentTok{#>  [3] "condoncollection"                                                      }
\CommentTok{#>  [4] "Taxonomy in Flux Checklist"                                            }
\CommentTok{#>  [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"     }
\CommentTok{#>  [6] "Bryophyte Collection - University of Washington Herbarium (WTU)"       }
\CommentTok{#>  [7] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"}
\CommentTok{#>  [8] "UWFC Ichthyology Collection"                                           }
\CommentTok{#>  [9] "Lichen Collection - University of Washington Herbarium (WTU)"          }
\CommentTok{#> [10] "UWBM Mammalogy Collection"}
\end{Highlighting}
\end{Shaded}

See also \texttt{datasets()} and \texttt{dataset\_suggest()} for
searching for datasets.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2637).

@@

@@

Networks, nodes and installations are at a higher level of organization
above datasets, but can be useful if you want to explore data from given
organizations. Here, we search for the first 10 GBIF networks, returning
just the title field.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit=}\DecValTok{10}\NormalTok{)$data$title}
\CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
\CommentTok{#>  [2] "Canadensys"                                                 }
\CommentTok{#>  [3] "Southwest Collections of Arthropods Network (SCAN)"         }
\CommentTok{#>  [4] "VertNet"                                                    }
\CommentTok{#>  [5] "Dryad"                                                      }
\CommentTok{#>  [6] "GBIF Network"                                               }
\CommentTok{#>  [7] "The Knowledge Network for Biocomplexity (KNB) "             }
\CommentTok{#>  [8] "Online Zoological Collections of Australian Museums (OZCAM)"}
\CommentTok{#>  [9] "Catalogue of Life"                                          }
\CommentTok{#> [10] "Ocean Biogeographic Information System (OBIS)"}
\end{Highlighting}
\end{Shaded}

@@

The GBIF taxonomic names API services are spread across five functions:

@@

@@

We can search the backbone taxonomy with the function
\texttt{name\_backbone()}. Here, we're searching for the name
\emph{Poa}, restricting to genera, and the family \emph{Poaceae}.

@@

@@

\begin{longtable}[c]{@{}rll@{}}
\toprule
key & canonicalName & rank\tabularnewline
\midrule
\endhead
3269133 & Pumilus & GENUS\tabularnewline
4407849 & Pumilinura & GENUS\tabularnewline
4323990 & Pumiliopsis & GENUS\tabularnewline
4324083 & Pumiliopes & GENUS\tabularnewline
4161281 & Pumilea & GENUS\tabularnewline
4312370 & Pumilopagurus & GENUS\tabularnewline
\bottomrule
\end{longtable}

@@

GBIF provides two ways to get occurrence data: through the
\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
\texttt{/occurrence/download} route (many functions, see below).
\texttt{occ\_search()} is the main function for the search route, and is
more appropriate for smaller data, while \texttt{occ\_download*()}
functions are more appropriate for larger data requests.

Large is of course a subjective term. When you hit a ``large dataset''
will depend primarily on the size of the your data request. GBIF imposes
for any given search a limit of 200,000 records in the search service,
after which point you can't download any more records for that search.
However, you can download more records for different searches.

We think the search service is still quite useful for many people even
given the 200,000 limit. For those that need more data, we have created
a similar interface in the \texttt{download\_*()} functions, that should
be easy to use. Users should take note that using the download service
has a few extra steps to get data into R, but is straight-forward.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
not do that in the occurrence search API. Thus, unfortunately, we
couldn't make the query interface exactly the same for both search and
download functions.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
\NormalTok{dat %>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
\CommentTok{#>       gbifID decimalLatitude decimalLongitude}
\CommentTok{#> 1  725767384        61.01005         24.41740}
\CommentTok{#> 2  725767447        59.82923         23.13550}
\CommentTok{#> 3  725767450        60.38505         25.17449}
\CommentTok{#> 4  725767513        68.37648         23.51963}
\CommentTok{#> 5  725767546        67.19203         24.85820}
\CommentTok{#> 6  725767579        60.21607         24.67412}
\CommentTok{#> 7  725767609        66.49260         25.70471}
\CommentTok{#> 8  725767645        61.36634         24.76218}
\CommentTok{#> 9  725767678        62.29174         27.96500}
\CommentTok{#> 10 725767681        60.28615         22.38489}
\CommentTok{#> ..       ...             ...              ...}
\end{Highlighting}
\end{Shaded}

@@

The downloaded dataset from GBIF is actually a Darwin Core Archive
(DwC-A), an internationally recognized biodiversity informatics standard
(\url{http://rs.tdwg.org/dwc/}). The DwC-A downloaded is a compressed
folder with a number of files, including metadata, citations for each of
the datasets included in the download, and the data itself, in separate
files for each dataset as well as one single \texttt{.txt} file. In
\texttt{occ\_download\_import()}, we simply fetch data from the
\texttt{.txt} file. If you want to dig into the metadata, citations,
etc., it is easily accessible from the folder on your computer.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrences - \texttt{occ\_count()}
\item
  Search for occurrences - \texttt{occ\_search()}
\item
  A simplified and optimized version of \texttt{occ\_search()} -
  \texttt{occ\_data()}
\item
  Get occurrences by occurrence identifier - \texttt{occ\_get()}
\item
  Get occurrence metadata - \texttt{occ\_metadata()}
\end{itemize}

@@

@@

The output of \texttt{occ\_search()} borrows the tidy
\texttt{data.frame} idea from the \texttt{dplyr} R package, so that no
matter how large the \texttt{data.frame}, the output is easily assessed
because only a few of the records (rows) are shown, only a few columns
are shown (with others shown in name only), and metadata is shown on top
of the \texttt{data.frame} to indicate data found and returned, media
records found, unique taxonomic hierarchies returned, and the query
executed.

@@

Get only those occurrences with spatial issues

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={R and Python clients for GBIF species occurrence data},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{R and Python clients for GBIF species occurrence data}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
    \author[boettig]{Carl Boettiger}
   \ead{carl(at)ropensci.org}

@@

Perhaps the most fundamental element in many fields of ecology is the
individual organism. The number of individuals of each species in a
given location forms the basis for many sub-fields of ecology and
evolution. Some research questions necessitate collecting new data,
while others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
(Brown, 1995; Beck et al., 2012).

@@

\section{Global Biodiversity Information
Facility}\label{global-biodiversity-information-facility}

There are many online services that collect and maintain specimen
records. However, Global Biodiversity Information Facility (hereafter,
GBIF, \href{http://www.gbif.org}{\url{http://www.gbif.org}}) is the
largest collection of biodiversity records globally, currently with 643
million records, 1.6 million taxa, 15,450 datasets from 780 publishers
(as of 2016-02-09). Many large biodiversity warehouses such as
iNaturalist
(\href{http://www.inaturalist.org}{\url{http://www.inaturalist.org}}),
VertNet (\href{http://vertnet.org}{\url{http://vertnet.org}}), and
USGS's Biodiversity Information Serving Our Nation (BISON;
\href{http://bison.usgs.ornl.gov}{\url{http://bison.usgs.ornl.gov}}) all
feed into GBIF.

The most important data organizational level in GBIF is the individual
record. In R using \texttt{rgbif} you'll recognize this as a row in the
data.frame that is returned from \texttt{occ\_search()} or
\texttt{occ\_data()}. Going upstream, each record is part of a dataset,
where each dataset is submitted by an organization, organizations are
organized into nodes, datasets are published through institutions (which
may be hosted at another organization), and a network is a group of
datasets (managed by GBIF).

Each record has some taxonomic name associated with it, which itself is
linked to a lot of other taxonomic data. GBIF maintains their own
taxonomic data.

All this organization matters because you can navigate through GBIF data
through all these organizational levels.

\section{The clients}\label{the-clients}

\subsection{The rgbif package}\label{the-rgbif-package}

Herein, we describe the \texttt{rgbif} software package (Chamberlain et
al.) for working with GBIF data in the R programming environment (R Core
Team, 2014). R is a widely used language in academia, as well as
non-profit and private sectors. Importantly, R makes it easy to execute
all steps of the research process, including data management, data
manipulation and cleaning, statistics, and visualization. Thus, an R
client for getting GBIF data is a powerful tool to facilitate
reproducible research.

@@

\subsection{The pygbif library}\label{the-pygbif-library}

\texttt{pygbif} (Chamberlain) is a Python library for working with GBIF
data in the Python programming environment ({\textbf{???}}). Python is a
general purpose programming language used widely in all sectors, and for
all parts of software development including server and client side use
cases. Python is used exclusively in some scientific disciplines (e.g.,
astronomy), and has partial usage in other disciplines. A Python client
for GBIF data is an important tool given the even wider usage of Python
than R, though maybe slightly less than R for ecology/biology.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pip install pygbif}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{import} \NormalTok{pygbif}
\end{Highlighting}
\end{Shaded}

The \texttt{pygbif} library is less mature and complete than the R
package. It also uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{pygbif} is developed publicly on GitHub at
\href{https://github.com/sckott/pygbif}{\url{https://github.com/sckott/pygbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{pygbif} can be
installed from \href{https://pypi.python.org/pypi/pygbif}{pypi}, the
distribution network for Python libraries.

\subsubsection{Library interfaces}\label{library-interfaces}

\texttt{rgbif} and \texttt{pygbif} are designed following the
\href{http://www.gbif.org/developer/summary}{GBIF Application
Programming Interface}, or API. The GBIF API has four major components:
registry, taxonomic names, occurrences, and maps. We also include
functions to interface with the OAI-PMH GBIF service; only dataset
information is available vis this service, however. We ignore maps in
both libraries as it is concerned with generating maps primarily for web
applications. Both libraries have a suite of functions dealing with each
of registry, taxonomic names, and occurrences - we'll go through each in
turn describing design and example usage.

\subsection{GBIF headers}\label{gbif-headers}

With each request \texttt{rgbif} and \texttt{pygbif} make to GBIF's API,
we send request headers that tell GBIF what client the request is coming
from, including what version of the library. This helps GBIF know what
proportion of requests are coming from which client, and therefore from
R vs.~Python; this information is helpful for GBIF in thinking about how
people are using GBIF data.

@@

Dataset information in general is available via the OAI-PMH service,
functions in \texttt{rgbif} prefixed with \texttt{gbif\_oai\_}, but not
available in \texttt{pygbif} yet.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{res <-}\StringTok{ }\KeywordTok{dataset_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"oregon"}\NormalTok{)}
\NormalTok{res$data$datasetTitle[}\DecValTok{1}\NormalTok{:}\DecValTok{10}\NormalTok{]}
\CommentTok{#>  [1] "UCDavis - Western USA - Monarch Butterflies - 1892-2005"                                                                         }
\CommentTok{#>  [2] "A geographic distribution database of Mononychellus mites (Acari: Tetranychidae) on cassava (Manihot esculenta)"                 }
\CommentTok{#>  [3] "Plantas Acuáticas de la Orinoquía Colombiana"                                                                                    }
\CommentTok{#>  [4] "USBombus, contemporary survey data of North American bumble bees (Hymenoptera,  Apidae, Bombus) distributed in the United States"}
\CommentTok{#>  [5] "Wool carder bees of the genus Anthidium in the Western Hemisphere"                                                               }
\CommentTok{#>  [6] "CM Birds Collection"                                                                                                             }
\CommentTok{#>  [7] "SDNHM Birds Collection"                                                                                                          }
\CommentTok{#>  [8] "University of British Columbia Herbarium (UBC) - Bryophytes Collection"                                                          }
\CommentTok{#>  [9] "University of British Columbia Herbarium (UBC) - Vascular Plant Collection"                                                      }
\CommentTok{#> [10] "Bryophyte Collection - University of Washington Herbarium (WTU)"}
\end{Highlighting}
\end{Shaded}

See also \texttt{datasets()} and \texttt{dataset\_suggest()} in
\texttt{rgbif} for searching for datasets.

In Python, we can similary search for datasets. Here, search for
datasets of type \texttt{OCCURRENCE}:

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
\NormalTok{registry.datasets(}\DataTypeTok{type}\NormalTok{=}\StringTok{"OCCURRENCE"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2737).

@@

And in Python, get metrics for the same dataset as above:

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
\NormalTok{registry.dataset_metrics(uuid=}\StringTok{'ec93a739-1681-4b04-b62f-3a687127a17f'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Networks, nodes and installations are at a higher level of organization
above datasets, but can be useful if you want to explore data from given
organizations. Here, in R we search for the first 10 GBIF networks,
returning just the title field.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{networks}\NormalTok{(}\DataTypeTok{limit =} \DecValTok{10}\NormalTok{)$data$title}
\CommentTok{#>  [1] "GBIF Backbone Sources"                                      }
\CommentTok{#>  [2] "Canadensys"                                                 }
\CommentTok{#>  [3] "Southwest Collections of Arthropods Network (SCAN)"         }
\CommentTok{#>  [4] "VertNet"                                                    }
\CommentTok{#>  [5] "Dryad"                                                      }
\CommentTok{#>  [6] "GBIF Network"                                               }
\CommentTok{#>  [7] "The Knowledge Network for Biocomplexity (KNB) "             }
\CommentTok{#>  [8] "Online Zoological Collections of Australian Museums (OZCAM)"}
\CommentTok{#>  [9] "Catalogue of Life"                                          }
\CommentTok{#> [10] "Ocean Biogeographic Information System (OBIS)"}
\end{Highlighting}
\end{Shaded}

And in Python:

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{registry}
\NormalTok{registry.networks(limit = }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The GBIF taxonomic names API services are spread across five functions
in \texttt{rgbif}:

@@

\texttt{pygbif} only has \texttt{name\_backbone()} and
\texttt{name\_suggest()} at this time.

@@

We can search the backbone taxonomy with the function
\texttt{name\_backbone()} in both R and Python clients. Here, we're
searching for the name \emph{Poa}, restricting to genera, and the family
\emph{Poaceae}, in R

@@

and in Python

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
\NormalTok{res = species.name_backbone(name=}\StringTok{'Poa'}\NormalTok{, rank=}\StringTok{'genus'}\NormalTok{, family=}\StringTok{'Poaceae'}\NormalTok{)}
\NormalTok{[ res[x] }\KeywordTok{for} \NormalTok{x in [}\StringTok{'usageKey'}\NormalTok{, }\StringTok{'kingdom'}\NormalTok{] ]}
\end{Highlighting}
\end{Shaded}

@@

\begin{longtable}[c]{@{}rll@{}}
\toprule
key & canonicalName & rank\tabularnewline
\midrule
\endhead
4848380 & Pumiliornis & GENUS\tabularnewline
2235094 & Pumilibranchipus & GENUS\tabularnewline
1795637 & Pumora & GENUS\tabularnewline
4581745 & Pumicia & GENUS\tabularnewline
4598790 & Pumilicopta & GENUS\tabularnewline
1593095 & Pumilomyia & GENUS\tabularnewline
\bottomrule
\end{longtable}

The same in Python

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{species}
\NormalTok{species.name_suggest(q=}\StringTok{'Pum'}\NormalTok{, limit = }\DecValTok{6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

GBIF provides two ways to get occurrence data: through the
\texttt{/occurrence/search} route (see \texttt{occ\_search} in
\texttt{rgbif}, or \texttt{occurrences.search} in \texttt{pygbif}), or
via the \texttt{/occurrence/download} route (many functions, see below).
\texttt{occ\_search()}/\texttt{occurrences.search} is the main function
for the search route, and is more appropriate for smaller data, while
the download functions are more appropriate for larger data requests.

GBIF imposes for any given search a limit of 200,000 records in the
search service, after which point you can't download any more records
for that search. However, you can download more records for different
searches.

We think the search service is still quite useful for many people even
given the 200,000 limit. For those that need more data, we have created
a similar interface in the download functions, that should be easy to
use. Users should take note that using the download service has a few
extra steps to get data into R, but is straight-forward.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you cannot
do that in the occurrence search API. Thus, unfortunately, we couldn't
make the query interface exactly the same for both search and download
functions.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
\NormalTok{dat %>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
\CommentTok{#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience}
\CommentTok{#> 1  725767384       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 2  725767447       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 3  725767450       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 4  725767513       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 5  725767546       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 6  725767579       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 7  725767609       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 8  725767645       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 9  725767678       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> 10 725767681       NA                         NA                 NA            NA          NA       NA}
\CommentTok{#> ..       ...      ...          ...           ...                ...           ...         ...      ...}
\CommentTok{#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),}
\CommentTok{#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted}
\CommentTok{#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),}
\CommentTok{#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),}
\end{Highlighting}
\end{Shaded}

In Python

\begin{Shaded}
\begin{Highlighting}[]
\CharTok{from} \NormalTok{pygbif }\CharTok{import} \NormalTok{occurrences }\CharTok{as} \NormalTok{occ}
\NormalTok{occ.download(}\StringTok{'taxonKey = 3119195'}\NormalTok{)}
\NormalTok{(res = occ.download_get(}\StringTok{"0000840-150615163101818"}\NormalTok{, overwrite = }\OtherTok{True}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

We don't have \texttt{pygbif} functionality at the moment for importing
data, but it's coming soon.

@@

The downloaded dataset from GBIF is a Darwin Core Archive (DwC-A), an
internationally recognized biodiversity informatics standard
(\href{http://rs.tdwg.org/dwc/}{\url{http://rs.tdwg.org/dwc/}}). The
DwC-A downloaded is a compressed folder with a number of files,
including metadata, citations for each of the datasets included in the
download, and the data itself, in separate files for each dataset as
well as one single \texttt{.txt} file. In
\texttt{rgbif::occ\_download\_import()}, we simply fetch data from the
\texttt{.txt} file. If you want to dig into the metadata, citations,
etc., it is easily accessible from the folder on your computer.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrences - \texttt{rgbif}:
  \texttt{occ\_count()} / \texttt{pygbif}: \texttt{occurrences.count}
\item
  Search for occurrences - \texttt{rgbif}: \texttt{occ\_search()} /
  \texttt{pygbif}: \texttt{occurrences.count}
\item
  A simplified and optimized version of \texttt{rgbif}:
  \texttt{occ\_search()} or \texttt{occ\_data()} / \texttt{pygbif}:
  \texttt{occurrences.count}
\item
  Get occurrences by occurrence identifier - \texttt{rgbif}:
  \texttt{occ\_get()} / \texttt{pygbif}: \texttt{occurrences.count}
\item
  Get occurrence metadata - \texttt{rgbif}: \texttt{occ\_metadata()} /
  \texttt{pygbif}: \texttt{occurrences.count}
\end{itemize}

@@

\textbf{R}

@@

The output of \texttt{occ\_search()} presents a compact
\texttt{data.frame} so that no matter how large the \texttt{data.frame},
the output is easily assessed because only a few of the records (rows)
are shown, only a few columns are shown (with others shown in name
only), and metadata is shown on top of the \texttt{data.frame} to
indicate data found and returned, media records found, unique taxonomic
hierarchies returned, and the query executed.

@@

Get only those occurrences with spatial issues. Spatial issues are a set
of issues that are returned in the \texttt{issues} field. They each
indicate something different about that record. For example, the issue
\texttt{COUNTRY\_COORDINATE\_MISMATCH} indicates that the interpreted
occurrence coordinates fall outside of the indicated country. You can
see how that might be useful when it comes to cleaning your data prior
to analysis/visualization.

@@

\textbf{Python}

The equivalent occurrence search workhorse in \texttt{pygibf} is
\texttt{occurrences.search}.
